---
title: "Resource"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
# Load the data
mydata <- read.table("./clean.csv", header=TRUE, sep=",")
source('visuals.R')
```


## Visuals

```{r pressure, echo=FALSE}
# Import the dataset
femDir <- mydata[mydata$director_gender == 1,]
ggplot(femDir, aes(x=startYear))+geom_freqpoly(aes(colour=director_gender), size=3)
mDir <- mydata[mydata$director_gender == 2,]
ggplot(mDir, aes(x=startYear))+geom_freqpoly(aes(colour=director_gender), size=3)
```

### Gender composition
```{r echo=FALSE}
v.ratings
```

This plot shows the relationship between gender composition of a film and its average viewer ratings.
Our predictions show that there is a slight positive correlation between a more male-dominated cast
and a slight negative correlation with a more female-dominated cast.

```{r echo=FALSE}
v.revenue
```

This plot shows the relationship between gender composition of a film and another, 
arguably more important metric of success, revenue.
The model for this scenario suggests that there is a positive correlation associated with high
percentages of both genders, with a male-dominated cast being about 3 times more highly correlated.